\chapter{Description logic}\label{chap:Desc}

\section{Introduction to description logic}

This chapter is based on Handbook on ontologies \cite{handbook} and Handbook of knowledge representation \cite{handbook2}.
\\ \\
The word ``ontology'' is used with different meanings in different communities.  \\
In philosophy, Aristotle in his Metaphysics defined \textbf{Ontology} as the study of attributes that belong to things because of their very nature.\\ Ontology focuses on the
nature and structure of things independently of any further considerations,
and even independently of their actual existence.\\ For example, it makes
perfect sense to study the Ontology of unicorns and other fictitious entities:
although they do not have actual existence, their nature and structure can be
described in terms of general categories and relations.
\\ \\
in Computer Science, we refer to an \textbf{ontology} as a special kind of information object or
computational artifact. Computational ontologies are a means to formally model the structure
of a system, that is the relevant entities and relations that emerge from its
observation, and which are useful to our purposes. An example of such a
system can be a company with all its employees and their interrelationships.
The ontology engineer analyzes relevant entities and organizes them into concepts
and relations, being represented, respectively, by unary and binary predicates.
\\ \\
\textbf{Description logics} (\textbf{DLs}) are a family of knowledge representation
languages that can be used to represent an ontology
in a structured and formally well-understood way. The ''description'' part of their name is based on how the important notions of the domain
are described by concept descriptions (unary predicates) and atomic roles (binary predicates). The ''logic'' part comes from their formal, logic-based semantics,
unlike some other methods of representation of ontologies, for example semantic networks. In \textbf{DLs}, ontology is represented by a knowledge base.
\begin{mydef} (Knowledge base)
A Knowledge base is an ordered pair ($\T,\A$) where $\T$ is the terminological part (\textbf{TBox}) and $\A$ is the assertional part (\textbf{ABox}). 
\end{mydef}
\begin{mydef} (GCI, TBox)
The terminological part of a knowledge base, \textbf{Tbox}, is a finite set of general concept inclusions.
A general concept inclusion (\textbf{GCI}) is a statement of the form $A \sqsubseteq B$ , where A and B are concepts. \\
The definition of a concept varies depending on the \textbf{DL}, for a definition in ALC, see \ref{ALCsyntax}.
\end{mydef}
\begin{myex} \label{TBoxEx}(GCI) 
Examples of \textbf{GCIs} in description logic ALC, and their equivalents in natural language (note that conventionally, concept names start with an uppercase letter, while individual and role names start with a lowercase letter) : \\
\begin{itemize}
\item $Human \sqsubseteq \forall hasChild.Human$ \\
''If someone is a human, then all their children are also human'' 
\item  $ Human  \sqcap  \neg  Female  \sqcap  (\exists  married.\top) \sqcap
(\forall  hasChild.(Doctor  \sqcup  Professor)) \sqsubseteq HappyMan $  \\
''A human who is not female, is married, and every child he has is either a doctor or a proffesor, is a happy man.'' \\
Note that $\top$ is a special concept, meaning ''everything'' - every individual is an instance of this concept. Similarily, there is a special concept $\bot$, or ''nothing'' - no individual is an instance of this concept. 

\item $Painter \equiv \exists hasPainted.Picture$ \\
$ A \equiv B$ is a shorter version of $(A \sqsubseteq B)  \land (B \sqsubseteq A)$, so this statement means
''If someone has painted a picture, they are a painter, and if someone is a painter, they have painted a picture''. 
\end{itemize}
\end{myex}
\begin{mydef} (ABoxEx)
\textbf{ABox} consists of 2 types of statements about individuals : concept assertions and role assertions. \\
Concept assertions have the form $a : C$ where a is an individual and C is a concept. \\
Role assertions have the form $a,b:R$ where a,b are individuals and R is a role.   
\end{mydef}
\begin{myex} 
\label{ABox}
Some examples of ABox statements, with no natural language translations as their meaning is obvious: 
\begin{itemize}
\item $Alice : Human$ 
\item $michelangelo : Painter \sqcap Sculptor$ 
\item $daVinci , monaLisa : hasPainted$
\end{itemize}
\end{myex}

\textbf{Interpretation} of description logics is done using sets. Although formal definitions vary based on the type of description logic, the general idea is that we create a set of objects, assign one object to every individual, interpret concepts as sets of objects and roles as sets of ordered doubles of objects. A formal definition can be seen in the next section, \ref{ALCsemantics}. 
\section{Selected description logics}

\subsection{$\mathcal{ALC}$}
In this thesis, we will be using a widely used description logic $\mathcal{ALC}$ and it's extensions. Although
$\mathcal{ALC}$ stands for ''Attributive concept Language with Complements''. It's one of the less expressive languages, for example, it can't express the concept '' someone who has 2 children ''.
You can see examples of TBox (\ref{TBoxEx}) and ABox (\ref{ABoxEx}) statements in $\mathcal{ALC}$ in the previous section, now it is time for formal definitions. 
\begin{mydef} 
\label{ALCsyntax}
\cite{handbook2}
(Syntax of $\mathcal{ALC}$ concepts and roles). Let $N_C$ be a set of concept names and $N_R$ be a set of
role names. The set of Concepts is the smallest set such that
\begin{enumerate}
\item $ \top, \bot,$ and every concept name $A \in N_C$ is an Concept,
\item  If C and D are Concepts and $r \in N_R$, then $C \sqcap D,\\ C \sqcup D , \neg C, \forall r.C,$ and $\exists r.C$ are Concepts.
\end{enumerate}
The set of roles is the set of role names $N_R$. (Note: this is not true for some simple extensions of $\ALC$, for example $\ALC _{(\cap)}$ - $\ALC$ with role intersection)  
\end{mydef}
The semantics of $\mathcal{ALC}$ (and of DLs in general) are given as interpretations. 
\begin{mydef}
\label{ALCsemantics}
\cite{handbook2} 
($\mathcal{ALC}$ semantics). An interpretation $\mathcal{I}$ = ($\Delta ^\mathcal{I}$, $\cdot ^\mathcal{I}$) consists of a nonempty
set $\Delta ^\mathcal{I}$, called the domain of $\mathcal{I}$, and a function $\cdot ^\mathcal{I}$ that maps every $\mathcal{ALC}$ Concept
to a subset of $\Delta ^\mathcal{I}$, and every $\mathcal{ALC}$ role to a subset of $\Delta ^\mathcal{I} \times \Delta ^\mathcal{I}$ such that, for all $\mathcal{ALC}$ Concepts C, D and all role names r:\\
\begin{tabular}{l l}
 $\top ^\mathcal{I} = \Delta^\mathcal{I}$ & $\bot ^\mathcal{I} = \emptyset $, \\
$(C \sqcap D)^\mathcal{I}$ & $= C^\mathcal{I} \cap D^\mathcal{I}$,\\
 $(C \sqcup D) ^\mathcal{I}$ & $= C^\mathcal{I} \cup D^\mathcal{I}$\\
 $\neg C^\mathcal{I}$ & $= \Delta^\mathcal{I} \backslash C^\mathcal{I}$, \\
 $ (\exists r.C)^\mathcal{I}$ & $= \{x \in \Delta ^\mathcal{I} | \exists y \in \Delta ^\mathcal{I} $ with  $\langle x, y \rangle \in r^\mathcal{I}$ and $y \in C^\mathcal{I}\}$,\\
 $(\forall r.C)^\mathcal{I}$ & $= \{x \in \Delta ^\mathcal{I} | \forall y \in \Delta ^\mathcal{I},$ if $\langle x, y \rangle \in  r^\mathcal{I},$ then  $y \in C^\mathcal{I}\}$.\\
\end{tabular}
\end{mydef}
You may notice that an interpretation does not have to comply with any TBox or ABox assertions.  If some assertion A is ''true'' for an interpretation $\mathcal{I}$, we say that $\mathcal{I}$ \textbf{satisfies} A.

\begin{mydef} (satisfiability)
Given an interpretation $\mathcal{I}$:
\begin{itemize}
\item  A GCI $A \sqsubseteq B$ is satisfied by $\mathcal{I}$ iff $A^\mathcal{I} \subset   B^\mathcal{I} $ .
\item  A concept assertion $a:C$ is satisfied by $\mathcal{I}$ iff $a^\mathcal{I} \in C^\mathcal{I} $ .
\item A role assertion $a,b:R$ is satisfied by $\mathcal{I}$ iff $ \langle a^mathcal{I} , b^mathcal{I} \rangle \in R^\mathcal{I} $
\end{itemize}
An assertion is satisfiable iff an interpretation exists that satisfies it.
\end{mydef}

\begin{mydef} (model, consistency)
An interpretation $\mathcal{I}$ is a model of a knowledge base $\K$ if it satisfies every assertion in the TBox and ABox of $\K$.  A knowledge base is consistent if at least one model of it exists.
\end{mydef}


An ontology can have multiple models, some less intuitive than other.

\begin{myex} (model) \\

\noindent Let there be a knowledge base $\mathcal{K} = \{ \mathcal{T} , \mathcal{A} \}$, 
$ \mathcal{T} $= \{ \\
\indent $A \sqsubseteq B$ \\
\indent $A \sqsubseteq	\exists r.C$ \\
\indent $C \sqsubseteq \forall r.D$\\
\}\\
$ \mathcal{A} $ = \{  \\
\indent a : A \\
\indent c,d : r \\
\} 
\\
One possible model of $\mathcal{K}$ would be $\mathcal{M}_1$ = \{ \\
\indent $\Delta ^\mathcal{I}$ = \{ $a_x, c_x, d_x$ \} \\
\indent $a^\mathcal{I} = a_x, c^\mathcal{I} = c_x, d^\mathcal{I} = d_x$\\
\indent $A^\mathcal{I} = \{a_x\}, B^\mathcal{I} = \{a_x\}, C^\mathcal{I} = \{c_x\}, D^\mathcal{I} = \{d_x\} $\\
\indent $r^\mathcal{I} = \{ \langle a_x,b_x\rangle , \langle a_x,c_X \rangle, \langle c_x,d_x \rangle \} $ \\
  		\} \\
But other, less intuitive models also exist, for example $\mathcal{M}_2$ and $\mathcal{M}_3$.
$\mathcal{M}_2$ = \{  \\
\indent$ \Delta ^\mathcal{I}$ = \{ $a_x, c_x, d_x, c_n$ \} \\ 
\indent$a^\mathcal{I} $ = $a_x, c^\mathcal{I} = c_x, d^\mathcal{I} = d_x$ \\
\indent$A^\mathcal{I}$ = $\{a_x\}, B^\mathcal{I} = \{a_x\}, C^\mathcal{I} = \{c_n\}, D^\mathcal{I} = \{ \} $ \\
\indent$r^\mathcal{I}$ = $\{ \langle a_x,b_x\rangle , \langle a_x,c_n \rangle, \langle c_x,d_x \rangle \} $
\} \\
$\mathcal{M}_3$ = \{ \\
\indent $\Delta ^\mathcal{I} = \{ i_x\}     $\\
 \indent $a^\mathcal{I} = i_x, c^\mathcal{I} = i_x, d^\mathcal{I} = i_x $\\
\indent $A^\mathcal{I} = B^\mathcal{I} = C^\mathcal{I} = D^\mathcal{I} = \{i_x\}$\\
\indent $r^\mathcal{I} = \{ \langle i_x, i_x \rangle \}$\\
\}
\end{myex}

\subsection{$\mathcal{SHIQ}$} 
\label{SHIQ}
A new description logic can be defined by adding some syntax to an existing description logic, increasing it's expressibility. As such, description logics are usually named as a list of abbreviations (TODO: abbreviations? acronyms? nenapada ma ako preformulovat tuto vetu) of the allowed expressions.\\  
The language $\mathcal{SHIQ}$ is one of the more expressive description logics. $\mathcal{S}$ is an abbreviation of ''$\mathcal{ALC}$ with transitive roles'' - meaning we can use any syntax from $\ALC$, and also declare a role to be transitive.
% meaning that for a transitive role R and objects (TODO? tak som ich pomenoval v definicii interpretacie, niesom si isty ci to je spravny nazov) (a,b,c) , if $\langle a,b\rangle \in R^\mathcal{I} $  and $\langle b,c\rangle \in R^\mathcal{I}$ , then $\langle a,c \rangle \in R^\mathcal{I}$. 
$\mathcal{R}$ stands for ''role hierarchy'', meaning we can define a role to be a subrole of another role. $\mathcal{I}$ stands for ''inverse roles'' (TODO spytat sa: wiki vravi ''inverse properties'', ale mne sa zda ze property je OWL termin a role je DL termin, a teraz vravime o DL a preto vraviet o properties bez definovania by bolo metuce) , meaning for every role R, there is an inverse role $R^-$. Lastly, $\mathcal{Q}$ stand for "qualified cardinality restriction" which allows us to specify the number of objects an object relates to by a role, for example we can say that a painter is someone who has painted at least 10 paintings.


\begin{myex} 
Some examples of TBox statements that can be made in $\mathcal{SHIQ}$ but not in $\mathcal{ALC}$:
\begin{itemize}
\item  $ Painter \sqsubseteq Human \sqcap \geq 10 hasPainted.Picture $
\item  $ Painting \sqsubseteq Picture \sqcap \exists hasPainted^-.\top$ \\ 
''A Painting is a picture that was painted.''
\item $isParent \sqsubseteq isAncestor$ \\
Meaning isParent is a subrole of hasAncestor - if someone is a parent of someone, they are also their ancestor.
\item $ transitive(isAncestor)$ \\
 The role isAncestor is transitive - if A is an ancestor of B and B is an ancestor of C, then A is an ancestor of C.
\end{itemize}
Note that the last 2 examples introduce  statements that are not general concept inclusion to the TBox. For this reason, some authors prefer defining a third part of the knowledge base, RBox, consisting of statements about roles.\\
\end{myex}


The definitions of syntax and semantics of $\mathcal{SHIQ}$ are similar to those of $\mathcal{ALC}$.
\begin{mydef} 
\cite{handbook} ($\mathcal{SHIQ}$ concept and role syntax) Let R be a set of role names, which is partitioned into a set $R_+$ of transitive roles and a set $R_p$ of normal roles. 
The set of all $\mathcal{SHIQ}$ roles is R $\cup \{r^- | r \in R\}$, where $r^-$ is called the inverse of the role r. \\
Let $N_C$ be a set of concept names. The set of $\mathcal{SHIQ}$ concepts is the smallest set such that:  
\begin{enumerate}
\item every concept $A \in N_C$ is a $\mathcal{SHIQ}$ concept.
\item if A and B are $\mathcal{SHIQ}$ concepts and r is a $\mathcal{SHIQ}$ role, then $A\sqcap B, A\sqcup B,  \neg A, \forall r.A,  and \exists r.A$ are $\mathcal{SHIQ}$ concepts.
\item if A is a $\mathcal{SHIQ}$ concept and r is a simple $\mathcal{SHIQ}$ role (simple role is neither transitive nor has a transitive subrole), and $n \in \mathbb{N}$, then $(\leq n r.A)$  and 
$(\geq n r.A)$ are $\mathcal{SHIQ}$ concepts.
\end{enumerate}
\end{mydef}
$\SHIQ$ semantics can be described as $\ALC$ semantics with same additions.

 \begin{mydef}($\mathcal{SHIQ}$ semantics) \cite{handbook}
in addition to definition \ref{ALC semantics}, for all  $p \in R$ and $r \in R_+$: \\
%\begin{itemize} 
 $\langle x,y \rangle \in p^\mathcal{I}$ iff  $\langle y,x \rangle \in (p^- )^\mathcal{I}$. \\
 if $\langle x,y \rangle \in r^\mathcal{I}$ and $ \langle y,z \rangle \in r^\mathcal{I}$ then $\langle x,z \rangle \in r^\mathcal{I}$.\\
 $(\leq n r.C)^\mathcal{I} = \{x \in \Delta ^\mathcal{I} | \# r^\mathcal{I} (x,C) \leq n\}$,\\
$(\geq n r.C)^\mathcal{I} = \{x \in \Delta ^\mathcal{I} | \# r^\mathcal{I} (x,C) \geq n\}$,\\

%\end{itemize}
where \#M  denotes the cardinality of the set M, and $r^\mathcal{I} (x,C) \mathrel{\mathop:}= \{y | \langle x,y \rangle \in r^\mathcal{I}$ and $y \in C^\mathcal{I}\}$.
\end{mydef}

$\SHIQ$ ABox and it's model are the same as in $\ALC$ (definition \ref{ALCAbox}). For the TBox, we have to add role subsumption axioms. 
Some authors define TBox as only containing GCIs, and use a new structure, RBox, to contain role inclusions. For this thesis, role inclusion axioms are a part of TBox.

\begin{mydef} \label{SHIQTBox} \cite{handbook} ($\SHIQ$ TBox) \\
A role inclusion axiom is of the form $r \sqsubseteq s$, where r,s
are roles. An interpretation $\I$ is a model of a role inclusion axiom  $r \sqsubseteq s$ if $r^{\I} \subseteq s^{\I}$.\\
$\I$ is a model of a TBox $\T$ if it it is a model of every role inclusion axiom and GCI axiom (definition \ref{ALCTBox}) in TBox.
\end{mydef}







\subsection{$\mathcal{ALCHO}$}
$\mathcal{ALC}$ with role hierarchy and Nominals.\\
\textbf{Nominals} are concepts with exacly one specific instance. For example, \{john\} is a concept with its only instance being the individual 'john'. \\
Nominals can be used to express enumerations, for example \cite{primer}: \\
   $Beatle \equiv \{john\} \sqcup  \{paul\} \sqcup  \{george\} \sqcup  \{ringo\} $\\
\\
\textbf{Role hirerarchy} was already described in section \ref{SHIQ} (example \ref{rolehierarchy}).
\begin{mydef} (Syntax and semantics of $\mathcal{ALCHO}$)\\
	Syntax of $\mathcal{ALCHO}$ is the syntax of $\mathcal{ALC}$ (definition \ref{ALCsyntax}), with \{a\} added to the set of concepts C for each individual 'a'.\\
	Similarly, semantics of $\mathcal{ALCHO}$ are the semantics of $\mathcal{ALC}$ (definition \ref{ALCsemantics}) with the addition of 
     $\{a\}^\mathcal{I}$ = $ \{ a^\mathcal{I} \} $ ,\\ 
     where 'a' is an individual and $ \mathcal{I}$ is the interpretation. \\
\end{mydef}
In $\ALCHO$, the definition of an ABox is the same for $\ALC$ (definition \ref{ALCAbox}) and the definition of a TBox is the same as for $\SHIQ$ (definition \ref{SHIQTBox}.
\subsection{$\mathcal{SROIQV(D)}$}
Jazyk koncludu, TODO asi ked bude hotova implementacia.
\section{Tableau algorithm}

In our algorithm, we heavily make use of tableau algorithm.
Tableau algorithm is a method of constructing a model of a knowledge base $\mathcal{K}$ if $\mathcal{K}$ is consistent, and stops if no model of $\mathcal{K}$ exists and thererfore $\K$ is inconsistent. \\ \indent
Tableau algorithm uses knowledge base in negation normal form (NNF), that is, every concept complement $\neg$ applies only to a concept name \cite{handbook2}.  Any $\mathcal{ALC}$ concept can be transformed to an equivalent concept in NNF by using de Morgan's laws and the duality between existential and universal restrictions ($\neg \exists r.C \equiv \forall r.{\neg C}$).
For example, the concept $\neg (\exists r.A \sqcap \forall s.B)$,
where A, B are concept names, can be transformed using de Morgan's laws to $\neg \exists r.C \sqcup \neg \forall s.B$, and this can then be transformed using the existential-universal duality into $(\forall r.\neg A) \sqcup (\exists s.\neg B)$.
\\ \\ 
\indent The idea behind the tableau algorithm for $\mathcal{K} = \{ \mathcal{T} , \mathcal{A} \} $is to start with the concrete situation described in $\mathcal{A}$ and expand based on what can be inferred from $\mathcal{T}$ and currently known ABox statements. This is done using something called completion graph, which is a graph where nodes represent individuals, directed edges represent relations between individuals, each node has a label containing concepts the individual belongs to, and each edge has a label consisting of the names of it's roles .

\begin{mydef} (Completion graph) \\
A completion graph is a pair ($\G,\L$), where $\G$ is a directed finite graph and $\L$ is a labeling function mapping each node from $\G$ to a set of concepts, and each edge to a set of roles. 
\end{mydef}


Tableau algorithm for $\K (\T ,\A )$ first creates a completion graph based on $\A$, and then expands it using tableau expansion rules.

\begin{mydef} \cite{handbook2} ($\ALC$ tableau expansion rules)
\begin{itemize}
\item $\sqcap$-rule: if 	 $C_1 \sqcap C_2 \in \L (x)$, x is not blocked (definition \ref{blocking}), and 
$\{C_1, C_2\} \not\subseteq  \L (x)$
then set $\L (x) = \L (x) \cup \{C_1, C_2\}$
\item $\sqcup$-rule: if 1. $C_1 \sqcup C_2 \in \L (x)$, x is not blocked, and 
$\{C_1, C_2\} \cap  \L (x) = \emptyset$,
then set $\L (x) = \L (x) \cup \{C\}$, for some $C \in \{C_1,C_2\}$
\item $\exists$-rule: if  $\exists r.C \in \L (x)$, x is not blocked, and
x has no r-successor y with $C \in \L (y)$,
then create a new node y with $\L (\langle x, y \rangle) = \{r\}$ and $\L (y) = \{C\}$.
\item $\forall$-rule: if $\forall r.C \in \L (x)$, x is not blocked, and
there is an r-successor y of x with $C \not\in \L (y)$
then set $\L(y) = \L(y) \cup \{C\}$.
\item $\sqsubseteq$-rule: if  $C_1 \sqsubseteq C_2 \in \T $, x is not blocked, and
$C_2 \sqcup $NNF($\neg C_1) \not\in \L (x)$,
then set $\L (x) = \L (x) \cup \{C_2 \sqcup$ NNF($\neg C_1$) \} 
\end{itemize}
Where $C, C_1, C_2$ are concepts, $r$ is role, $\T$ is TBox, and NNF is normal negation form.
\end{mydef}

Why are we checking whether nodes are blocked? And what are blocked nodes? \\
If we removed ''x is not blocked'' from all expansion rules, then the algorithm could generate an infinite graph using the $\exists$-rule, for example for 
$\K =(\T ,\A), \T = \{C \sqsubseteq \exists r.C \}, \A=\{a:C\}$ the algorithm would create a node $a_2$ with $\langle a,a_2 \rangle :r$, then a node $a_3$ with $\langle a_2,a_3\rangle :r$ and so on.
To guarantee termination, we define node blocking.
\begin{mydef} \label{ancestor} (ancestor)
In completion graph $\mathcal{CG} = (G, \L), G=(N,E)$, nodes $x,y \in N$, and edge $\langle x,y \rangle \in E$, then y is a successor of x and x is an ancestor of y. \\
If $\L (\langle x,y \rangle) = r$, then y is an r-successor of x. \\
Ancestry is transitive (if node x is an ancestor of node y and y is an an ancestor of node z, then x is is an ancestor of z).

\end{mydef}
\begin{mydef} \label{blocking}(blocking)
A node x is blocked if it has an ancestor y that is either blocked, or $\L (x) \subseteq \L (y)$ .
\end{mydef}

And now we can show the pseudocode of (nondeterministic) tableau algorithm.

\begin{algorithm}[t!]
    \begin{algorithmic}[1]
		    \Function{TableauAlg}{ Knowledge base $\K$ =($\T ,\A$)}
			\State Convert $\T$ to NNF.
			\State Completion graph $\CG =(\G,\L) , \G=(\Nodes,\Edges), \Nodes = \Edges = \L =\emptyset$
			\For{Each individual $i \in \A$}
				\State $\Nodes = \Nodes \cup n_i$
				\For{Each assertional axiom $(i : C) \in \A$}
					\State $\L (n_i) = \L (n_i) \cup C$
				\EndFor
				
				\For{Each assertional axiom $(i,x : r) \in \A$}
					\State $\Edges = \Edges \cup \langle n_i,n_x \rangle$
					\State $\L (\langle n_i,n_x \rangle) = \L (\langle n_i,n_x \rangle)  \cup r$
				\EndFor  
			\EndFor
			\While{A tableau expansion rule can be applied on $\CG$}
				\State Apply a rule on $CG$
				\If{Clash exists in $\CG$}
					\State \Return ''$\K$ is inconsistent''  (assuming algorithm undeterministically always picks the 'correct' decision in $\sqcup$-rule if one exists)
				\EndIf
			\EndWhile
			\State \Return ''$\K$ is consistent''
			\EndFunction

    \end{algorithmic}
\end{algorithm}

	\begin{mydef} (clash)\\
		In completion graph $\CG$=$( (\Nodes, \Edges),\L)$ , if there is a node $n \in \Nodes$ and a concept C where $\{ C , \neg C \} \in \L (n)$, then $\CG$ has a clash.
	
\end{mydef}
If the algorithm cannot apply any expansion rules, then $\K$ is consistent and $\CG$ is its model, or a finite part of an infinite model due to blocking. Our abduction algorithm only needs this finite part.\\
 The tableau algorithm has 2 sources of nondeterminism : picking which rule to apply, and how to apply the $\sqcup$-rule. The order of rule applications doesn't matter for the sake of determining consistency, but the application of the $\sqcup$-rule does, as one choice can lead to a clash while the other not. \\ 
If a clash happens in a deterministic implementation, we can backtrack to a previous $\sqcup$ decision with an unexplored choice and continue from there. If there are no unexplored $\sqcup$ decisions and we find a clash, $\K$ is inconsistent (no model exists).     
If there is no decision with an unexplored possibility we can backtrack to, we can declare $\mathcal{K}$ to be inconsistent. 
I think it would be useful to show an example of tableau completion graph:
\\
obrazok odstraneny, nejake problemy s hboxom a nemoznostou lomit riadok.

\begin{myex}
\begin{align*}
 \K &= \left\{
     \begin{aligned}
      \mathsf{Student} & \sqsubseteq \exists \mathsf{owns}.\mathsf{Textbook} \\       
		\mathsf{Textbook} & \sqsubseteq \mathsf{Book} \\
 		\mathsf{tom} & \colon \mathsf{Student}
     \end{aligned}
   \right\}\\
 \end{align*}

 \medskip
 \begin{center}
 \begin{tikzpicture}[-latex ,auto ,node distance =0.5 cm and 0.5 cm,on grid , semithick, align=left]
	\node[draw,circle,inner sep=2pt,fill] (a) at (0,0) {};
	\node (a1) [above left=of a] {$\mathsf{tom}$}; 
	\node (a2) at (0,-2) {\\$\{\mathsf{Student}$\\$ \neg \mathsf{Student} \sqcup \exists \mathsf{owns}.\mathsf{Textbook}$\\$ \exists \mathsf{owns}.\mathsf{Textbook}$\\$ \neg \mathsf{Textbook} \sqcup \mathsf{Book},$\\ $\neg \mathsf{Textbook} \}$} ; 
	\node[draw,circle,inner sep=2pt,fill] (b) at (6,0) {}; 
	\path (a) edge [bend right = -15] node[above =0.15 cm] {$\mathsf{owns}$} (b); 
	\node (b1) [above right=of b] {$\mathsf{b_1}$}; 
	\node (b2) at (6,-2) {$\{\mathsf{Textbook}$,\\$ \neg \mathsf{Student}  \sqcup \exists \mathsf{owns}.\mathsf{Textbook} $\\ $\neg \mathsf{Student} $\\ $\neg \mathsf{Textbook} \sqcup \mathsf{Book}$\\$\mathsf{Book}$}; 

	\end{tikzpicture}  
 \end{center}
In this case, the tableau algorithm stops without a clash, therefore the knowledge base is consistent.
\end{myex}


%TODO casova zlozitost, rozsirenie tableau pre ostatne spomenute deskripcne logiky.
%\section{OWL (mozno)}
%	TODO: definovat SHOIN
%\section{Použitia  (možno, spomenúť rôzne databázy)}

